---
title: "Homework 6"
author: "Lindsey Carlson"
date: "2/23/2022"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

# Simulating and Fitting Data Distributions  
This exercise teaches students how to compare a histogram of continuous (or integer) data to the probability density functions for different statistical distributions.

### Question 1
Run Sample Code    
```{r, ECHO = FALSE}

library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation

# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)


p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)


p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute

meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat


expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
 
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
 
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
 
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```





### Question 2
Try it with your own data - Data provided by LTER

```{r, ECHO = TRUE}
z <- read.table("lpp_hglvl_data.csv",header=TRUE,sep=",")
z <- z[z$MEAN_HG_CONC>0,]
str(z)
summary(z)

library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation


p1 <- ggplot(data=z, aes(x=MEAN_HG_CONC, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)


p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

normPars <- fitdistr(z$MEAN_HG_CONC,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute

meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$MEAN_HG_CONC),len=length(z$MEAN_HG_CONC))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$MEAN_HG_CONC), args = list(mean = meanML, sd = sdML))
 p1 + stat
 
#red shows a normal distribution to the curve


expoPars <- fitdistr(z$MEAN_HG_CONC,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$MEAN_HG_CONC), args = list(rate=rateML))
 p1 + stat + stat2
 
# blue shows an exponential distribution

 
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$MEAN_HG_CONC), args = list(min=min(z$MEAN_HG_CONC), max=max(z$MEAN_HG_CONC)))
 p1 + stat + stat2 + stat3
 
# dark green shows uniform distribution
 
gammaPars <- fitdistr(z$MEAN_HG_CONC,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$MEAN_HG_CONC), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
 
#brown displays gamma distribution
 
pSpecial <- ggplot(data=z, aes(x=MEAN_HG_CONC/(max(MEAN_HG_CONC + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$MEAN_HG_CONC/max(z$MEAN_HG_CONC + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$MEAN_HG_CONC), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```


### Question 3 
Find best-fitting distribution

The **gamma curve** fits the data the best.




### Question 4  
Simulate a new dataset 
```{r}
z_sim <- rnorm(n=43,mean=normPars$estimate["mean"])
z_sim <- data.frame(1:43,z_sim)
names(z_sim) <- list("ID","myVar")
z_sim <- z_sim[z_sim$myVar>0,]
str(z_sim)
summary(z_sim$myVar)


p1sim <- ggplot(data=z_sim, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1sim)


p1sim <-  p1sim +  geom_density(linetype="dotted",size=0.75)
print(p1sim)

normParssim <- fitdistr(z_sim$myVar,"normal")
print(normParssim)
str(normParssim)
normParssim$estimate["mean"] # note structure of getting a named attribute

meanMLsim <- normParssim$estimate["mean"]
sdMLsim <- normParssim$estimate["sd"]

xvalsim <- seq(0,max(z_sim$myVar),len=length(z_sim$myVar))

#plotting the best fit curve for original dataset - gamma 
gammaParssim <- fitdistr(z_sim$myVar,"gamma")
shapeMLsim <- gammaPars$estimate["shape"]
rateMLsim <- gammaPars$estimate["rate"]

stat4sim <- stat_function(aes(x = xvalsim, y = ..y..), fun = dgamma, colour="brown", n = length(z_sim$myVar), args = list(shape=shapeMLsim, rate=rateMLsim))
p1sim + stat4sim


#reprinting original data historgram with gamma curve 
p1+stat4
```

The two histograms do not show similar gamma distributions.  The model is not large enough to simulate realistic data to match my original measurements.  With more samples, the gamma distribution may not be the best fit, and thus the simulation may be more realistic.


[Home Page](index.html)