---
title: "Homework 10"
author: "Lindsey Carlson"
date: "3/30/2022"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

## For Loops and Randomization Tests 
*Worked on some parts of the code with Stephen Peters Collaer*

### Question 1 - Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the matrix. Finally, use return(counter) for the output.

``` {r}

##################################################
# FUNCTION: num_zero
# purpose: count number of zeros 
# input: x vectors of numeric
# output: count of number of zeros
#------------------------------------------------- 
num_zeros <- function(vector1 = sample(0:3, size = 1000, replace = TRUE)){
  counter <- 0
  for(i in 1:length(vector1)){
    if(vector1[i]== 0){
      counter <- counter + 1
    }
  }
  return(counter)
} 
 
num_zeros()

``` 

### Question 2 - Use subsetting instead of a loop to rewrite the function as a single line of code.

```{r}
##################################################
# FUNCTION: num_zero_sub
# purpose: count number of zeros 
# input: x vectors of numeric
# output: count of number of zeros
#------------------------------------------------- 
num_zeros_sub <- function(vector1 = sample(0:3, size = 1000, replace = TRUE)){
 length(vector1[vector1 == 0])
}
  
num_zeros_sub()

```


### Question 3 - Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

``` {r}
##################################################
# FUNCTION: mat_product
# purpose: multiply the row number x by column number
# input: number of rows and columns in a matrix
# output: product of the row number and column number 
#------------------------------------------------- 

mat_product <- function(nrow=4, ncol=4){
  mat <- matrix(nrow=4, ncol=4)
  for(i in 1:nrow){
    for(j in 1:ncol){
      mat[i,j] <- i*j
    }
  }
  print(mat)
}

mat_product()

``` 

### Question 4 - In the next few lectures, you will learn how to do a randomization test on your data. We will complete some of the steps today to practice calling custom functions within a for loop. Use the code from the March 31st lecture (Randomization Tests) to complete the following steps:

*a. Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.*
```{r}
group <- c(rep("A", 100), rep("B", 100), rep("C", 100))
A <- rnorm(n=100, mean = 1, sd = 0.1)
B <- rnorm(n=100, mean = 2, sd = 0.1)
C <- rnorm(n=100, mean = 3, sd = 0.1)

response <- c(A, B, C)
d_frame <- data.frame(group, response)

print(d_frame)
```

*b. Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.*
``` {r}
library(dplyr)

##################################################
# FUNCTION: reshuffle
# purpose: get means of reshuffled data
# input: groups of data (A, B, C)
# output: vector length of 3 with means of each group
#------------------------------------------------- 
reshuffle <- function(df = d_frame, response = response){
  colnames(d_frame) <- c('group', 'response')
  new_shuff <- sample(1:nrow(d_frame), replace=FALSE)
    d_frame <- data.frame(group, new_shuff)
  sum_mean <- d_frame %>% group_by(group) %>% summarise(mean = mean(new_shuff))
      return(sum_mean)
}

reshuffle()

```

*c. Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.* 

```{r}
library(tidyr)

results <- data.frame()
  
for(i in 1:100){
  shuff <- sample(1:nrow(d_frame), replace=FALSE)
    d_frame_shuff <- data.frame(group, shuff)
  sum_mean <- d_frame_shuff %>% group_by(group) %>%    
    summarise(mean = mean(shuff))
  df <- sum_mean %>% pivot_wider(names_from = group, values_from = mean)
  final_df <- cbind(i, df)
  
  results <- rbind(results, final_df)
}

print(results)

```

*d. Use qplot() to create a histogram of the means for each reshuffled group. Or, if you want a challenge, use ggplot() to overlay all 3 histograms in the same figure. How do the distributions of reshuffled means compare to the original means?*

```{r}
library(ggplot2)

mean_plotA <- qplot(x=A, data = results,
                   geom = "histogram",
                   xlab = "Mean of Group A",
                   ylab = "Count",
                   color=I("black"),
                   fill=I("Steelblue3"))

  print(mean_plotA) 
  
mean_plotB <- qplot(x=B, data = results,
                   geom = "histogram",
                   xlab = "Mean of Group B",
                   ylab = "Count",
                   color=I("black"),
                   fill=I("Steelblue3"))

  print(mean_plotB) 

mean_plotC <- qplot(x=C, data = results,
                   geom = "histogram",
                   xlab = "Mean of Group C",
                   ylab = "Count",
                   color=I("black"),
                   fill=I("Steelblue3"))

  print(mean_plotC) 

```
`
[Home Page](index.html)